Figures for 3-D printed oxygen devices

### 2015-08-13 publishing/response-to-reviewers.md

•	Page 4 – Why was 6 hrs chosen for cell culture? It would seem that the authors could demonstrate a more robust change in expression over the concentrations tested if the experiment was extended to a longer time point, especially considering only the 0% group showed a significant difference in VEGFA expression.

**Authors' rebuttal:** 
VEGFA is under the transcriptional activity of HIF-1alpha. HIF protein peaks at 4 hr, and so we looked at a time point slightly delayed. By longer time points, the HIF response is actually suppressed, and many transcripts are not increased with longer time points. See Rexius-Hall et al.s' work.

•	The authors specify concentrations of 21%, 10%, 5%, and 0% were evaluated, however in their Figure 2b, the concentrations appear closer to 13% (not 10%) and 7% (not 5%). Variation of 3% can be considerable in physiological environments and it is unclear if the authors recognize this variation. It is also amplified by the lack of error bars for each group.

**Authors' rebuttal:**
The data in the original submission was analyzed with a two point calibration. This method correlates intensites to oxygen concentration assuming a linear Stern-Volmer relationship. While standard practice, the two-point calibration suffers from increasing inaacuracy as points are measured further away from the calibration points.

### 2015-08-13 publishing/response-to-reviewers.md

•	Page 4 – Why was 6 hrs chosen for cell culture? It would seem that the authors could demonstrate a more robust change in expression over the concentrations tested if the experiment was extended to a longer time point, especially considering only the 0% group showed a significant difference in VEGFA expression.

**Authors' rebuttal:** 
VEGFA is under the transcriptional activity of HIF-1alpha. HIF protein peaks at 4 hr, and so we looked at a time point slightly delayed. By longer time points, the HIF response is actually suppressed, and many transcripts are not increased with longer time points. See Rexius-Hall et al.s' work.

•	The authors specify concentrations of 21%, 10%, 5%, and 0% were evaluated, however in their Figure 2b, the concentrations appear closer to 13% (not 10%) and 7% (not 5%). Variation of 3% can be considerable in physiological environments and it is unclear if the authors recognize this variation. It is also amplified by the lack of error bars for each group.

**Authors' rebuttal:**
The data in the original submission was analyzed with a two point calibration. This method correlates intensities to oxygen concentration assuming a linear Stern-Volmer relationship. While standard practice, the two-point calibration suffers from increasing inaccuracy as points are measured further away from the calibration points. Characterization data was re-taken with 5 calibration points and analyzed with a two-site Stern-Volmer model to increase accuracy of analyzed data. A new plot replaces figure 2 with error bars.

@@

•	Is sterilization an issue for this device, considering it will be in close contact with biological components?

**Authors' rebuttal:**
We did not observe any contamination.

### 2015-08-13 publishing/response-to-reviewers.md

•	The authors specify concentrations of 21%, 10%, 5%, and 0% were evaluated, however in their Figure 2b, the concentrations appear closer to 13% (not 10%) and 7% (not 5%). Variation of 3% can be considerable in physiological environments and it is unclear if the authors recognize this variation. It is also amplified by the lack of error bars for each group.

**Authors' rebuttal:**
The data in the original submission was analyzed with a two point calibration. This method correlates intensities to oxygen concentration assuming a linear Stern-Volmer relationship. While standard practice, the two-point calibration suffers from increasing inaccuracy as points are measured further away from the calibration points. Characterization data was re-taken with 5 calibration points and analyzed with a two-site Stern-Volmer model to increase accuracy of analyzed data. A new plot replaces figure 2 and includes error bars.

•	Is sterilization an issue for this device, considering it will be in close contact with biological components?

**Authors' rebuttal:**
Sterilization was preformed by spraying the device with 70% ethonal and leaving it in a biosaftey hood with UV lights overnight. We did not observe any contamination. This information was added to the manuscript.

### 2015-08-13 publishing/latex/revision/revision.tex
updating manuscript

Oxygen was measured with PtOEPK (Pt(II) Octaethylporphine ketone) planar sensors that were fixed to the bottom of a 24-well plate.
Sensors were prepared as discussed previously \cite{Sinkala2010,Rexius-Hall2014}. 
Briefly, polystyrene was dissolved in toluene to form a solution to which PtOEPK (Frontier Scientific) was mixed in before spin coating to a thin layer and allowing the toluene to evaporate.
The intensity of the sensor was measured with via fluorescence microscopy and correlated to the concentration of oxygen with Stern-Volmer analysis.

@@

Human lung adenocarcinoma (A549) cells were seeded at 30,000 cells per well in a 24-well plate.
Cells were cultured in DMEM supplemented with 10\% FBS.
After reaching $\sim$70\% confluency, the 24-well insert was placed in the plate and the gases were perfused through the device in the same scheme as in the oxygen characterization.
Each row of six wells experienced either 0\%, 5\%, 10\% or 21\% \ce{O2}, with 5\% \ce{CO2} and balanced nitrogen.
The assembly was placed in an incubator at $37\,^{\circ}$C for six hours.

@@

Following six hours of oxygen modulation, the 24-well oxygen control insert was removed and culture media was aspirated.
Cells in each well were washed twice with PBS, and 300 $\mu$L of a lysis solution consisting of 10 $\mu$L of $\beta$-mercaptoethanol (Sigma Aldrich) per 1 mL of lysis buffer (PureLink RNA Mini Kit; Invitrogen) was added to each well.
The lysate from each well was collected as a sample.
RNA was extracted according to the PureLink RNA Mini Kit manufacturer’s instructions.

@@

% Results and Discussion can be combined.
\section*{Results}
\subsection*{Oxygen control}
Oxygen control in the 24-well insert device was quantified with a platinum-based (PtOEPK) planar oxygen sensor placed at the bottom of the well, as shown in Fig.~\ref{fig2}.
The device reaches steady state in 30 minutes and can then hold the oxygen level near 0\% \ce{O2} indefinitely, while 95\% \ce{N2}, 5\% \ce{CO2} is pumped through the device. 
A different gas condition can be used in each separate iteration of the 6-well unit allowing 6 technical replicates of up to 4 different gas conditions in one 24-well plate (Fig.~\ref{fig2}). There was little variation in the oxygen levels measured in each of the 6 wells controlled by a single inlet, demonstrating the distribution network worked effectively.

\begin{figure}[h]
\caption{
{\bf Oxygen Characterization.} 
(A) Time course data of oxygen being evacuated from the culture area as 0\% oxygen gas is perfused through the device.
Error bars are the standard deviation N=3.
(B) Four oxygen conditions are demonstrated in a 24-well plate.
Each 6-well row of the plate can be controlled independently.  
Mean N=3 error bars not included.
}
\label{fig2}
\end{figure}

>>>>

Oxygen was measured with PtOEPK (Pt(II) Octaethylporphine ketone) planar sensors that were fixed to the bottom of a 24-well plate.
Sensors were prepared as discussed previously \cite{Sinkala2010,Rexius-Hall2014}. 
Briefly, polystyrene was dissolved in toluene to form a solution to which PtOEPK (Frontier Scientific) was mixed in before spin coating to a thin layer and allowing the toluene to evaporate.
Five calibration points were fitted to a two-site model\cite{Carraway1991,Wang2014} of the Stern-Volmer relationship which was used to correlate sensor intensities with oxygen concentration.

@@

The assembled device was sterilized by spraying with 70\% ethanol solution and placed in a biosaftey hood with UV lights overnight.
Human lung adenocarcinoma (A549) cells were seeded at 30,000 cells per well in a 24-well plate.
Cells were cultured in DMEM supplemented with 10\% FBS.
After reaching $\sim$70\% confluency, the 24-well insert was placed in the plate and the gases were perfused through the device in the same scheme as in the oxygen characterization.
Each row of six wells experienced either 0\%, 5\%, 10\% or 21\% \ce{O2}, with 5\% \ce{CO2} and balanced nitrogen.
The assembly was placed in an incubator at $37\,^{\circ}$C for six hours.

@@

Following six hours of oxygen modulation, the 24-well oxygen control insert was removed and culture media was aspirated.
No contamination of the culture was observed.
Cells in each well were washed twice with PBS, and 300 $\mu$L of a lysis solution consisting of 10 $\mu$L of $\beta$-mercaptoethanol (Sigma Aldrich) per 1 mL of lysis buffer (PureLink RNA Mini Kit; Invitrogen) was added to each well.
The lysate from each well was collected as a sample.
RNA was extracted according to the PureLink RNA Mini Kit manufacturer’s instructions.

@@

% Results and Discussion can be combined.
\section*{Results}
\subsection*{Oxygen control}
Oxygen control in the 24-well insert device was quantified with a platinum-based (PtOEPK) planar oxygen sensor placed at the bottom of the well, as shown in Fig.~\ref{fig2}.
The device reaches steady state in 30 minutes and can then hold the oxygen level near 0\% \ce{O2} indefinitely, while 95\% \ce{N2}, 5\% \ce{CO2} is pumped through the device.
A different gas condition can be used in each separate iteration of the 6-well unit allowing 6 technical replicates of up to 4 different gas conditions in one 24-well plate.
There was little variation in the oxygen levels measured in each of the 6 wells controlled by a single inlet, demonstrating the distribution network worked effectively.

\begin{figure}[h]
\caption{
{\bf Oxygen Characterization.} 
Time course data of four oxygen conditions are demonstrated in a 24-well plate.
Each 6-well row of the plate can be controlled independently.  
Error bars are the standard deviation N=3.
}
\label{fig2}
\end{figure}
